---
title: "EllenHutti_FinalProject"
author: "Ellen Hutti"
date: "11/18/2020"
output: html_document
---

## Library
```{r}
library(tidyverse)
library(dplyr)
library(haven)
library(table1)
library(odds.n.ends)
library(survey)
library(lmtest)
library(aod)
```

## Load data
```{r}
NISTEEN <- read_csv("https://raw.githubusercontent.com/ehutti/ADA-Final-Project/main/NISTEEN18.csv")
```

## Check HPV vaccine variable
```{r}
# Check variable class
class(NISTEEN$HPV_binary)
table(NISTEEN$HPV_binary)

# Make outcome variable
NISTEEN <- NISTEEN %>% 
  mutate(HPV_binary = 
           case_when(HPVI_ANY %in% c("No") ~ 0, 
                     HPVI_ANY %in% c("Yes") ~ 1),
         HPV_binary = factor(HPV_binary, 0:1, c("No", "Yes")))
```

## Create table 1
```{r}
# Add labels to the variables
label(NISTEEN$HPVI_ANY)<-"Received HPV vaccine"
label(NISTEEN$HPVI_RECOM)<-"HPV vaccine recommended by doctor"
label(NISTEEN$AGE)<-"Age in years"
label(NISTEEN$CEN_REG)<-"Geographic region"
label(NISTEEN$EDUC1)<-"Mother's education attainment"
label(NISTEEN$RACEETHK)<-"Race or ethnicity"
label(NISTEEN$SEX)<-"Sex"
label(NISTEEN$FACILITY)<-"Type of healthcare facility"
label(NISTEEN$INS_STAT2_I)<-"Type of insurance"
label(NISTEEN$INCQ298A)<-"Income"

# Create table 1
table1(~AGE + SEX + RACEETHK + INCQ298A + INS_STAT2_I + EDUC1 | HPVI_ANY, NISTEEN)
```

## Construct complex survey design
```{r}
options( survey.lonely.psu = "adjust" )

nis_design <- 
    svydesign(
        id = ~ SEQNUMT , 
        strata = ~ STRATUM , 
        weights = ~ RDDWT_C , 
        data = subset( NISTEEN , RDDWT_C > 0 ) 
    ) 
```

## Survey weighted univariate model
# Mother's education
```{r}
educglm <- 
    svyglm( 
        HPV_binary ~ EDUC1 , 
        nis_design ,
        family=binomial()
    )

summary( educglm )

OReduc<-exp(cbind(OR = coef(educglm), confint(educglm)))
OReduc

# Children of mother's with a high school diploma or equivalent are 0.77 (95% CI, 0.69-0.86) times as likely to receive the HPV vaccine compared to children of mother's with a college degree.
# Children of mother's with less than a high school diploma or equivalent are 0.85 (95% CI, 0.74-0.97) times as likely to receive the HPV vaccine compared to children of mother's with a college degree.
# Children of mother's with some college are 0.83 (95% CI, 0.75-0.91) times as likely to receive the HPV vaccine compared to children of mother's with a college degree.
```

# Insurance status
```{r}
insurglm <- 
    svyglm( 
        HPV_binary ~ INS_STAT2_I , 
        nis_design ,
        family=binomial()
    )

summary( educglm )

ORinsur<-exp(cbind(OR = coef(insurglm), confint(insurglm)))
ORinsur

# Those with other insurance are 0.84 (95% CI, 0.69-1.03) times as likely to receive the HPV vaccine compared to those with Medicaid insurance.
# Those with private insurance are 1.06 (95% CI, 0.93-1.20) times as likely to receive the HPV vaccine compared to those with Medicaid insurance.
# Those with no insurance are 0.65 (95% CI, 0.50-0.86) times as likely to receive the HPV vaccine compared to those with Medicaid insurance.
```

## Survey weighted ultivariate models
# Mother's education
```{r}
# Multivariate model for mother's education attainment adjusted for race or ethnicity. Race or ethnicity was chosen to adjust for based on the DAG created for this model. 
adjeducglm <- 
    svyglm( 
        HPV_binary ~ EDUC1 + RACEETHK, 
        nis_design ,
        family=binomial()
    )

summary( adjeducglm )

adjOReduc<-exp(cbind(OR = coef(adjeducglm), confint(adjeducglm)))
adjOReduc

# Children of mother's with a high school diploma or equivalent are 0.75 (95% CI, 0.67-0.84) times as likely to receive the HPV vaccine compared to children of mother's with a college degree after adjusting for race or ethnicity.
# Children of mother's with less than a high school diploma or equivalent are 0.81 (95% CI, 0.70-0.94) times as likely to receive the HPV vaccine compared to children of mother's with a college degree after adjusting for race or ethnicity.
# Children of mother's with some college are 0.82 (95% CI, 0.74-0.89) times as likely to receive the HPV vaccine compared to children of mother's with a college degree after adjusting for race or ethnicity.
```

# Insurance status
```{r}
# Multivariate model for insurance status adjusted for income, race or ethnicity, and mother's education attainment. Variables for adjustment were chosen based on the DAG created for this model. 
adjinsurglm <- 
    svyglm( 
        HPV_binary ~ INS_STAT2_I + INCQ298A + RACEETHK + EDUC1, 
        nis_design ,
        family=binomial()
    )

summary( adjinsurglm )

adjORinsur<-exp(cbind(OR = coef(adjinsurglm), confint(adjinsurglm)))
adjORinsur

# Those with other insurance are 0.86 (95% CI, 0.69-1.07) times as likely to receive the HPV vaccine compared to those with Medicaid insurance after adjusting for income, race, and mother's education attainment.
# Those with private insurance are 1.04 (95% CI, 0.87-1.23) times as likely to receive the HPV vaccine compared to those with Medicaid insurance after adjusting for income, race, and mother's education attainment.
# Those with no insurance are 0.72 (95% CI, 0.54-0.96) times as likely to receive the HPV vaccine compared to those with Medicaid insurance after adjusting for income, race, and mother's education attainment.
```

## Plot Cook’s Distance to assess influential observations
``` {r}
# Plot of Cook's distance to identify top 5 influential observations for adjusted education model
plot(adjeducglm, which=4, id.n=5, col="red") 

# Plot of Cook's distance to identify top 5 influential observations for adjusted insurance model
plot(adjinsurglm, which=4, id.n=5, col="red") 

# No concerning Cook's distances found for either model
```

## Test model significance
# Education attainment model
```{r}
# Use Wald test to assess significance of mother's education attainment
wald.test(b = coef(adjmineducglm), Sigma = vcov(adjmineducglm), Terms = 2:4)

# Results of Wald test (X2=33.6, p<0.01) suggest that the overall model effects of mother's education attainment are statistically significant.
```

# Insurance status model
```{r}
# Use Wald test to assess significance of mother's education attainment
wald.test(b = coef(adjinsurglm), Sigma = vcov(adjinsurglm), Terms = 2:4)

# Results of Wald test (X2=8.2, p<0.05) suggest that the overall model effects of insurance status are statistically significant.
```
